---
title: "HW3_econ"
author: "Arash Barmas"
date: "2/14/2019"
output:
  word_document: default
  html_document: default
---


```{r}
library(mvtnorm)
options(digits=4)
mu.x = 0.01; sig.x = 0.25; mu.y = 0.05; sig.y = 0.15

rho.xy = 0.99
sig.xy = rho.xy*sig.x*sig.y
Sigma.xy = matrix(c(sig.x^2, sig.xy, sig.xy, sig.y^2), 2, 2, byrow=TRUE)

n = 100
set.seed(123)
xy.vals = rmvnorm(n, mean=c(mu.x, mu.y), sigma=Sigma.xy) 

plot(xy.vals[,1], xy.vals[,2],pch = 20 ,cex=1.5, col="blue", xlab="x", ylab="y")
title("Bivariate normal: rho = 0.99")
abline(h=mu.y, v=mu.x)
segments(x0=0, y0=min(xy.vals[,2]), x1=0, y1=0, col="red")
segments(x0=min(xy.vals[,1]), y0=0, x1=0, y1=0, col="red")
pmvnorm(lower=c(-Inf, -Inf), upper=c(0, 0), mean=c(mu.x, mu.y), sigma=Sigma.xy)
```
As we can see in the plot, a strong positive linear association exits, and the joint probability, Pr(X=< 0, Y =< 0) is 0.3691 with error approximately = 0.


```{r}
rho.xy = 0.9
sig.xy = rho.xy*sig.x*sig.y
Sigma.xy = matrix(c(sig.x^2, sig.xy, sig.xy, sig.y^2), 2, 2, byrow=TRUE)

n = 100
set.seed(123)
xy.vals = rmvnorm(n, mean=c(mu.x, mu.y), sigma=Sigma.xy) 

plot(xy.vals[,1], xy.vals[,2],pch = 20 ,cex=1.5, col="blue", xlab="x", ylab="y")
title("Bivariate normal: rho = 0.9")
abline(h=mu.y, v=mu.x)
segments(x0=0, y0=min(xy.vals[,2]), x1=0, y1=0, col="red")
segments(x0=min(xy.vals[,1]), y0=0, x1=0, y1=0, col="red")
pmvnorm(lower=c(-Inf, -Inf), upper=c(0, 0), mean=c(mu.x, mu.y), sigma=Sigma.xy)

```
As we can see in the plot, a positive linear relation exists, but it's not as strong as the first one, and the joint probability, Pr(X=< 0, Y =< 0) is 0.3421 with error approximately = 0.

```{r}

rho.xy = 0.5
sig.xy = rho.xy*sig.x*sig.y
Sigma.xy = matrix(c(sig.x^2, sig.xy, sig.xy, sig.y^2), 2, 2, byrow=TRUE)

n = 100
set.seed(123)
xy.vals = rmvnorm(n, mean=c(mu.x, mu.y), sigma=Sigma.xy) 

plot(xy.vals[,1], xy.vals[,2],pch = 20 ,cex=1.5, col="blue", xlab="x", ylab="y")
title("Bivariate normal: rho = 0.5")
abline(h=mu.y, v=mu.x)
segments(x0=0, y0=min(xy.vals[,2]), x1=0, y1=0, col="red")
segments(x0=min(xy.vals[,1]), y0=0, x1=0, y1=0, col="red")
pmvnorm(lower=c(-Inf, -Inf), upper=c(0, 0), mean=c(mu.x, mu.y), sigma=Sigma.xy)

```
There is a noticeable but comparatively weak positive linear correlation between X and Y. The joint probability, Pr(X=< 0, Y =< 0) is 0.2574 with error approximately = 0.

```{r}

rho.xy = 0
sig.xy = rho.xy*sig.x*sig.y
Sigma.xy = matrix(c(sig.x^2, sig.xy, sig.xy, sig.y^2), 2, 2, byrow=TRUE)

n = 100
set.seed(123)
xy.vals = rmvnorm(n, mean=c(mu.x, mu.y), sigma=Sigma.xy) 

plot(xy.vals[,1], xy.vals[,2],pch = 20 ,cex=1.5, col="blue", xlab="x", ylab="y")
title("Bivariate normal: rho = 0")
abline(h=mu.y, v=mu.x)
segments(x0=0, y0=min(xy.vals[,2]), x1=0, y1=0, col="red")
segments(x0=min(xy.vals[,1]), y0=0, x1=0, y1=0, col="red")
pmvnorm(lower=c(-Inf, -Inf), upper=c(0, 0), mean=c(mu.x, mu.y), sigma=Sigma.xy)

```
There seems to be no linear dependence between X and Y. The joint probability, Pr(X=< 0, Y =< 0) is 0.1788 with error approximately = 0.


```{r}
rho.xy = -0.9
sig.xy = rho.xy*sig.x*sig.y
Sigma.xy = matrix(c(sig.x^2, sig.xy, sig.xy, sig.y^2), 2, 2, byrow=TRUE)

n = 100
set.seed(123)
xy.vals = rmvnorm(n, mean=c(mu.x, mu.y), sigma=Sigma.xy) 

plot(xy.vals[,1], xy.vals[,2],pch = 20 ,cex=1.5, col="blue", xlab="x", ylab="y")
title("Bivariate normal: rho = -0.9")
abline(h=mu.y, v=mu.x)
segments(x0=0, y0=min(xy.vals[,2]), x1=0, y1=0, col="red")
segments(x0=min(xy.vals[,1]), y0=0, x1=0, y1=0, col="red")
pmvnorm(lower=c(-Inf, -Inf), upper=c(0, 0), mean=c(mu.x, mu.y), sigma=Sigma.xy)
```
There is an extremely strong negative correlation between X and Y. The joint probability, Pr(X=< 0, Y =< 0) is 0.02025 with error approximately = 0.
